### 2022-03-17 paper/localpsf.tex
restructuring paper

\subsection{Motivation: inverse problems with highly informative data}
\label{sec:PDE_hessian_motivation}

@@

%In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
Our work builds upon a history of ``product-convolution`` approximations---weighted sums of convolution operators with spatially varying weights. A single convolution operator may be either high rank or low rank, so product-convolution approximations can (where applicable) approximate high rank operators with a small number of convolution operators. Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}.
For background, we recommend reading the following papers: \cite{PRODCONVGOOD}. We specifically build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i$ \cite{PRODCONVLIST}. A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}. With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate product-convolution approximation. In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}. But even with adaptive refinement, many matrix vector products with $\Aop$ are required. In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\phi_i$ are computed with each matrix vector product.

Our method of estimating the support of the impulse responses was inspired by resolution analysis in seismic imaging \cite{RESOLUTION}. In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_p$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $p$. If the width of $\phi_p$ is ten meters, then features of the parameter near $p$ can be accurately inferred from data if those features are larger than roughly ten meters in size. In \cite{RESOLUTION}, the width of $\phi_p$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion. In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions. Our method estimates the support of $\phi_p$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.

@@

@@

In this section we present the three main concepts that our method is based on. The method itself will be presented in Section \ref{sec:method}. First, we define moments of the impulse responses, $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}). Our method will use these moments extensively. Second, we present ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{eq:mean_and_covariance_estimation}). Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches. Third, we describe a method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}). Our method will interpolate impulse responses using this approximation.

@@

The quantity $\tau$ is a parameter that must be chosen by the user of the algorithm. The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$. However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$. We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results. The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.

>>>>

\subsection{Motivation: PDE constrained inverse problems}
\label{sec:PDE_hessian_motivation}

@@

%In general, product-convolution approximations take the form shown in \eqref{eq:product_convolution}, but $\convkernel_i$ and $w_i$ may be arbitrary functions.
Our work builds upon a history of ``product-convolution`` approximations---weighted sums of convolution operators with spatially varying weights. A single convolution operator may be either high rank or low rank, so product-convolution approximations can (where applicable) approximate high rank operators with a small number of convolution operators. Product-convolution approximations have been used in a wide variety of fields going back several decades \cite{PRODCONVLIST}. Often product-convolution methods are known as ``point spread function'' (PSF) methods, particularly in applied fields such as optics and image processing. There, the ``operator'' $\Aop$ is a physical process of distortion and blurring that occurs as rays of light pass through an optical system like a telescope or microscope. However, ``PSF'' is also sometimes used as an umbrella term to refer to other approximation methods that focus on interpolating impulse responses or impulse response features, including CUR-based low rank approximation \cite{NAGYPSFCUR}.
For background on product-convolution and PSF methods, we recommend reading the following papers: \cite{PRODCONVGOOD}.

We specifically build upon the class of product convolution approximations in which the convolution kernels are shifted impulse responses of $\Aop$ associated with a collection of points $x_i$ \cite{PRODCONVLIST}. A popular choice for methods in this class is to choose the points $x_i$ to be nodes in a regular grid, and interpolate the convolution operators with piecewise linear interpolation \cite{NAGY}, splines \cite{PRODCONVSPLINES}, or radial basis functions \cite{PRODCONVRBF}. With a regular grid, a large number of points $x_i$ is typically required to achieve an accurate approximation. In our previous work, we reduced the computational cost by starting with a coarse grid of points $x_i$, then adaptively refining the grid in the regions where the error in the approximation is large \cite{PRODCONVMYPAPER}. But even with adaptive refinement, many operator actions of $\Aop$ are required.

In this paper, rather than using adaptive refinement, we instead reduce the computational cost by picking points $x_i$ such that many $\phi_{x_i}$ are computed with each operator action of $\Aop$. This requires us to estimate the supports of the impulse responses, $\phi_x$, \emph{before} we compute them. We were inspired by resolution analysis in seismic imaging \cite{RESOLUTION}. In resolution analysis, $\Aop$ is the Hessian for a seismic inverse problem, and the width of $\phi_x$ is used to estimate of the minimum length scale on which features of the parameter can be inferred near the point $x$. If the width of $\phi_x$ is ten meters, then features of the parameter near $x$ can be accurately inferred from data if those features are larger than roughly ten meters in size. In \cite{RESOLUTION}, the width of $\phi_x$ is estimated to be the local autocorrelation length of the function $\Aop^T \zeta$ near $p$, where $\zeta$ is a random noise funtion. In this paper, rather than probing $\Aop^T$ with random noise functions, we use polynomial functions. Our method estimates the support of $\phi_x$ more accurately and reliably than resolution analysis, but our method requires that $\Aop$ has a positive integral kernel, while the resolution analysis method does not have this requirement.

@@

MOMENTS PERMIT IMPROVED INTERPOLATION. IMPROVED INTERPOLATION REQUIRES H-MATRIX. H-MATRIX WITH IMPROVED INTERPOLATION ALSO SOLVES BOUNDARY ARTIFACT PROBLEM. ALSO DOESNT REQUIRE REGULAR GRID. ALSO SCALABLE TO LARGE NUMBER OF IMPULSE RESPONSES. FFT HAS FAST MATVEC BUT NOT OTHER OPTIONS.

@@

In this section we present three main concepts that our method is based on. The method itself will be presented in Section \ref{sec:method}. First, we define moments of the impulse responses, $\phi_x$, and show how these moments can be computed efficiently (Section \ref{sec:intromoments}). Our method will use these moments extensively. Second, we present ellipsoid shaped a-priori estimates for the supports of the impulse responses (Section \ref{eq:mean_and_covariance_estimation}). Our method will use these ellipsoid estimates to choose sets of impulse responses that can be computed in batches. Third, we describe a method of approximating impulse responses by other nearby impulse responses, which we call ``normalized local mean displacement invariance'' (Section \ref{sec:local_mean_displacement_invariance}). Our method will interpolate impulse responses using this approximation.

@@

The quantity $\tau$ is a parameter that must be chosen by the user. The larger $\tau$ is, the larger the ellipsoid $E_x$ is, and the more conservative the estimate is for the support of $\impulseresponse_x$. However, in Section \ref{sec:sample_point_selection} we will see that the cost of our algorithm depends on how many non-overlapping ellipsoids $E_x$ we can ``pack'' in the domain $\Omega$ (more ellipsoids $\implies$ lower cost), and choosing a larger value of $\tau$ means that fewer ellipsoids will fit in $\Omega$. We find that $\tau=3$ yields a reasonable balance between these competing interests, and use $\tau=3$ in our numerical results. The fraction of the ``mass'' of $\impulseresponse_x$ residing outside of $E_x$ is less than $1/\tau^2$ by Chebyshev's inequality, though this bound is conservative and typically far less mass resides in this region.
